Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-15 concurrency/CONCUR 2023/paper.tex
WIP

For the purpose of reasoning about lock-coupling style (also known as hand-over-hand locking), we specifically consider traversing the tree using this locking method. To accomplish this, we must maintain at least one lock during traversal between nodes. This requires obtaining the lock for the next node before releasing the lock for the current node. Therefore, we utilize the \lstinline{pn} struct containing two \lstinline{node_t} type nodes: the \lstinline{p} field representing the current node, and the \lstinline{n} field denoting the next node.

@@

For many concurrent data structures, the ideal correctness condition requires that the data structure behaves identically to a sequential implementation, even when accessed concurrently by multiple threads. The concurrent behavior of search structures is specified using \emph{atomic triples}~\cite{tada}. These take the form of $\forall a.\ \left\langle \texttt{P}_l\ |\ \texttt{P}_p(a) \right\rangle\ \texttt{c}\ \left\langle \texttt{Q}_l\ |\ \texttt{Q}_p(a)\right\rangle$, where $\texttt{P}_l$ and $\texttt{Q}_l$ are \emph{local} preconditions and postconditions, akin to a standard Hoare triple, while $\texttt{P}_p$ and $\texttt{Q}_p$ are \emph{public} preconditions and postconditions, parameterized by an abstract value $a$ of the shared data structure. This triple suggests that if $\texttt{P}_l$ holds true before a call and $\texttt{P}_p$ is true for some value of $a$ in a shared state (e.g. a global invariant), then $\texttt{P}_p$ will continue to be true for some (possibly different) value of $a$ until the \emph{linearization point} of $\texttt{c}$, at which point $\texttt{Q}_p$ will become true atomically for the same value $a$ (and $\texttt{Q}_l$ will be true after $\texttt{c}$ ends). For instance, the specification
$\forall s.\ \left\langle \texttt{is\_stack}\ p\ |\ \texttt{stack}\ s\right\rangle\ \texttt{push}(v)\ \left\langle \texttt{is\_stack}\ p\ |\ \texttt{stack}\ (v::s)\right\rangle$
expresses the fact that the push operation of a concurrent stack correctly implements the behavior of a sequential push, atomically transitioning from particular stack $s$ to $v::s$ at some point during its execution.

@@

The particulars of $\nodeboxrep$ and $\treerep$ might differ between implementations, but at their core, $\nodeboxrep$ encompasses the resources tied to each client thread, while $\treerep$ acts as an abstract representation containing ghost state that connects the data structure's concrete state to its abstract counterpart. Specifically, $\nodeboxrep(\texttt{pn->n})$ functions as a pointer to both the node's lock and lock-invariant assertions, representing a protected resource through an arbitrary separation logic predicate. Acquiring this lock enables a thread to access further elements within the data structure. As previously mentioned, the \lstinline{pn} struct comprises two nodes of type \lstinline{node_t}: the $\texttt{p}$ field, which signifies the parent node, and the $\texttt{n}$ field, representing the current node.

@@

\than{FIX ME: These things above didn't represent hand-over-hand locking instance, since I tried to avoid talking about lock invariant.}
\than{I feel somewhat perplexed while attempting to articulate the situation mentioned above. Can someone help me fix or explain it more?}

@@

\section{Verified Data Structures}

\wm{I think this is the wrong organization. The proofs of traverse should be presented in the previous section, where we describe traverse. This section should show findNext and insert/lookup for specific data structure(s), and then describe how we combine them into top-level specs.}

>>>>

%For the purpose of reasoning about lock-coupling style (also known as hand-over-hand locking), we specifically consider traversing the tree using this locking method. To accomplish this, we must maintain at least one lock during traversal between nodes. This requires obtaining the lock for the next node before releasing the lock for the current node. Therefore, we utilize the \lstinline{pn} struct containing two \lstinline{node_t} type nodes: the \lstinline{p} field representing the current node, and the \lstinline{n} field denoting the next node.

@@

For many concurrent data structures, the ideal correctness condition requires that the data structure behaves identically to a sequential implementation, even when accessed concurrently by multiple threads. The concurrent behavior of search structures is specified using \emph{atomic triples}~\cite{tada}. These take the form of $\forall a.\ \left\langle \texttt{P}_l\ |\ \texttt{P}_p(a) \right\rangle\ \texttt{c}\ \left\langle \texttt{Q}_l\ |\ \texttt{Q}_p(a)\right\rangle$, where $\texttt{P}_l$ and $\texttt{Q}_l$ are \emph{local} preconditions and postconditions, akin to a standard Hoare triple, while $\texttt{P}_p$ and $\texttt{Q}_p$ are \emph{public} preconditions and postconditions, parameterized by an abstract value $a$ of the shared data structure. This triple suggests that if $\texttt{P}_l$ holds true before a call and $\texttt{P}_p$ is true for some value of $a$ in a shared state (e.g. a global invariant), then $\texttt{P}_p$ will continue to be true for some (possibly different) value of $a$ until the \emph{linearization point} of $\texttt{c}$, at which point $\texttt{Q}_p$ will become true atomically for the same value $a$ (and $\texttt{Q}_l$ will be true after $\texttt{c}$ ends). For instance, the specification
$\forall s.\ \left\langle \texttt{is\_stack}\ p\ |\ \texttt{stack}\ s\right\rangle\ \texttt{push}(v)\ \left\langle \texttt{is\_stack}\ p\ |\ \texttt{stack}\ (v::s)\right\rangle$
expresses the fact that the push operation of a concurrent stack correctly implements the behavior of a sequential push, atomically transitioning from particular stack $s$ to $v::s$ at some point during its execution. The stack itself is treated as a publicly owned resource, and can only be accessed and modified atomically by threads holding the \texttt{is\_stack} assertion.

To demonstrate the proof of an implementation that meets the atomic specification, let's take the example of the stack specification mentioned above. This specification assert that the implementation \texttt{push} appears to execute atomically, accessing and updating the state of the data structure without exposing any intermediate states. The local preconditions and postconditions, \texttt{is\_stack}, contain the points-to and lock-invariant assertions necessary for a thread, while the public preconditions and postconditions, \texttt{stack}, involve the ghost state that defines the overall state of the stack. Ghost state is a type of program state that aids in the verification process without affecting the program's runtime behavior, and by referring to an arbitrary ghost state identifier, it is possible to connect local and public conditions. To show that the implementation of \texttt{push}(v) satisfies these atomic specifications, two things must be demonstrated. Firstly, the implementation can execute safely without owning any piece of public assertion $\texttt{stack}$, only accessing it automatically and restoring it up until the linearization point. Secondly, at some linearization point, the implementation transforms the current stack $s$ into $v::s$ atomically, which satisfies the public postcondition $\texttt{stack}(v::s)$.

@@

The particulars of $\nodeboxrep$ and $\treerep$ might differ between implementations, but at their core, $\nodeboxrep$ encompasses the resources tied to each client thread, while $\treerep$ acts as an abstract representation containing ghost state that connects the data structure's concrete state to its abstract counterpart and is treated as an atomically accessed shared resource. Specifically, $\nodeboxrep(\texttt{pn->n})$ functions as a pointer to both the node's lock and lock-invariant assertions, representing a protected resource through an arbitrary separation logic predicate. Acquiring this lock enables a thread to access further elements within the data structure. As previously mentioned, the \lstinline{pn} struct comprises two nodes of type \lstinline{node_t}: the $\texttt{p}$ field, which signifies the parent node, and the $\texttt{n}$ field, representing the current node.

@@

%\than{FIX ME: These things above didn't represent hand-over-hand locking instance, since I tried to avoid talking about lock invariant.}
\than{I feel somewhat perplexed while attempting to articulate the situation mentioned above. Can someone help me fix or explain it more?}

@@

\section{Verified Data Structures}
\subsection{Lock-coupling Template}

\wm{I think this is the wrong organization. The proofs of traverse should be presented in the previous section, where we describe traverse. This section should show findNext and insert/lookup for specific data structure(s), and then describe how we combine them into top-level specs.}
\subsection{Give-up Template}
